---
title: "thesis_results"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
```

```{r}
source("Simulation_functions.R")
source("crossOverFunctions.R")
```

### CI interval clarification plot:
```{r}
observed_cn_dataframe_50 = obs_cn(n=10000, 
       read_length=100, 
       coverage=30,
       flank_seg_length=1000,
       cur_seg_length=50,
       actual_flank_seg_cn=2,
       actual_cur_seg_cn=1,
       fragment_size=600,
       rd_model =  "start_in",
       vaf_model = "split_read")

observed_cn_dataframe_100 = obs_cn(n=10000, 
       read_length=100, 
       coverage=30,
       flank_seg_length=1000,
       cur_seg_length=100,
       actual_flank_seg_cn=2,
       actual_cur_seg_cn=1,
       fragment_size=600,
       rd_model =  "start_in",
       vaf_model = "split_read")

observed_cn_dataframe_150 = obs_cn(n=10000, 
       read_length=100, 
       coverage=30,
       flank_seg_length=1000,
       cur_seg_length=150,
       actual_flank_seg_cn=2,
       actual_cur_seg_cn=1,
       fragment_size=600,
       rd_model =  "start_in",
       vaf_model = "split_read")

group_by(observed_cn_dataframe_150, model) %>% 
summarise(estimate_ci_width(cn_estimate))

observed_cn_dataframe_200 = obs_cn(n=10000, 
       read_length=100, 
       coverage=30,
       flank_seg_length=1000,
       cur_seg_length=200,
       actual_flank_seg_cn=2,
       actual_cur_seg_cn=1,
       fragment_size=600,
       rd_model =  "start_in",
       vaf_model = "split_read")

observed_cn_dataframe = bind_rows(observed_cn_dataframe_50, observed_cn_dataframe_100, observed_cn_dataframe_150, observed_cn_dataframe_200)
```

```{r}
working_dataframe = observed_cn_dataframe
working_dataframe$cur_seg_length = as.factor(working_dataframe$cur_seg_length)

# Compute CI estimate intervals of each model grouped by current segment length
group_by(working_dataframe, model, cur_seg_length) %>%
  summarise(estimate_ci_width(cn_estimate)) %>%
  view()

#group_by(working_dataframe, model, cur_seg_length) %>%
ggplot(working_dataframe, aes(x=cur_seg_length, 
           y=cn_estimate, 
           #group=model, 
           color=model)) +
  #geom_crossbar()
  #geom_errorbar(group=50, ymin=1 - 0.45, ymax=1 + 0.45)
  #geom_point(size=0.1) 
  geom_boxplot() +
  theme_bw() +
  #ggtitle("Interquantile range of estimated copy-number values \n from select current segment length sizes") +
  labs(x="Current segment length (b)", y="Copy-number estimate")
  #geom_errorbar(group="start_in", ymin=1 - 0.45, ymax=1 + 0.45)
  #geom_point(x=150, y=1) +
  #geom_errorbar(group=150, ymin=0.99 - 0.53, ymax=0.99 + 0.53, color="cyan", width = 0.5) +
  #geom_point(x=150, y=0.99, color="cyan") +
  #title="Interquantile range of estimated copy-number values from select current segment length sizes"
  

group_by(observed_cn_dataframe_50, model) %>%
summarise(median(cn_estimate))

view(working_dataframe)
```

### Flanking and current segment copy-number CI interval visualisation with corresponding VAF values.
In deletion cases for short current segment lengths (100 b for this plot), and any flanking segment length, VAF model performs better than RD model. Note that CI interval of inferred copy-numbers with SV-VAF < 0.5 is an acceptable range.  
```{r}
#' SV-VAF values for various flanking and current segment copy number values
#' Further confirms that SV-VAF model performs best when flanking and current CN values are closely related 
vaf_flank_cur_cn_sim = sim_permutations(
    cur_seg_length = 100,
    flank_seg_length = seq(100, 1000, 100),
    actual_flank_seg_cn = seq(0, 4, 0.2),
    actual_cur_seg_cn = seq(0, 4, 0.2),
    fragment_size = 600,
    rd_model = "start_in",
    vaf_model = "read_pair") %>%
      filter(model == "vaf read_pair") 

# This plot indicates that the vaf values less than 0.5 are more or less consistant with the assumption that in deletion cases SV-VAF model always perform better in short current segments. 
vaf_flank_cur_cn_sim = vaf_flank_cur_cn_sim %>%
  rename(flanking_segment_CN = actual_flank_seg_cn)
vaf_flank_cur_cn_sim = vaf_flank_cur_cn_sim %>%
  rename(current_segment_CN = actual_cur_seg_cn)
ggplot(vaf_flank_cur_cn_sim) +
  aes(x=vaf, y=ci_width, colour=model, shape=flanking_segment_CN > current_segment_CN) +
  geom_point(color = "blue") +
  scale_y_continuous(breaks=seq(0, 6, 1), limits = c(0, 6)) +
  geom_smooth(level = 0) +
  labs(colour="Model", 
       x = "SV-VAF",
       y = "Confidence Interval Width") +
  theme_bw()
```
## Hypothesis 
RD-based model performs better in amplification cases, where flanking segment copy-number is lower than current segment copy-number, when current segment length is $>50$, regardless of other parameter values. 
In amplification cases where flanking segment copy-number is lower than current segment copy-number, performance of RD-based model is always better (i.e., CI interval width of inferred copy-number is a lower value) regardless of the value of other parameters with the exception where current segment length is $\le50$, SV-VAF based model performed better. 
```{r}
# Takes about 2 hours RUN LATER
shortSegmentSim = sim_permutations(
  read_length = 150,
  cur_seg_length = seq(10, 500, 10),
  flank_seg_length = seq(50, 1000, 25),
  actual_flank_seg_cn = seq(1, 20, 1),
  actual_cur_seg_cn = seq(0.5, 4, 0.5),
  fragment_size = 600,
  coverage = 60,
  rd_model = "start_in",
  vaf_model = "read_pair")
```

```{r}
# The higher the log value the more similar the performance of two models are. 
# Higher log values = more similar performance
# DO NOT RUN AGAIN *** FINAL VALUES
lm2DF = sim_permutations(
  read_length = c(100, 150),
  cur_seg_length = seq(250, 400, 50),
  flank_seg_length = seq(1200, 1350, 50),
  actual_flank_seg_cn = c(1, 2, 3, 4, 30, 80),
  actual_cur_seg_cn = c(1, 2, 3, 4, 55, 70),
  fragment_size = c(400, 500, 600),
  coverage = c(30, 40, 60),
  rd_model = "start_in",
  vaf_model = "read_pair")

lm2_modelDF = lm2DF %>%
  combineCI(rd_model = "rd start_in", vaf_model = "vaf read_pair") %>%
  # Higher log values: higher similarity in performance between models
  mutate(lnCI_difference = -log(abs(vaf_ci_width - rd_ci_width))) 
# crossOverDF( rd_model = "rd start_in", 
  #              vaf_model = "vaf read_pair") %>%
  # mutate(lnCI_difference = -log(abs(vaf_ci_width - rd_ci_width)))
lm2DF_model1 = lm(lnCI_difference ~ cur_seg_length, lm2_modelDF)
summary(lm2DF_model1)
plot(lm2DF_model1)
lm2DF_model2 = lm(lnCI_difference ~ cur_seg_length + flank_seg_length + actual_flank_seg_cn + actual_cur_seg_cn + coverage + read_length + fragment_size, data = lm2_modelDF)
summary(lm2DF_model2)
# Check performance of two linear models
anova(lm2DF_model1, lm2DF_model2)
# Explore this interaction: cur_seg_length:flank_seg_length:actual_flank_seg_cn, -7.525e-08, to further investigate how these parameters are negatively correlated to the difference between model performance. 
lm2DF_model3 = lm(lnCI_difference ~ cur_seg_length * flank_seg_length * actual_flank_seg_cn * actual_cur_seg_cn * coverage * read_length * fragment_size, data = lm2_modelDF)
summary(lm2DF_model3)
anova(lm2DF_model2, lm2DF_model3)
# Linear model on a dataframe that only consist of similar performance between models to investigate which parameter(s) contirbute most to two models being similar.
someDF2 = lm(lnCI_difference ~ cur_seg_length + flank_seg_length + actual_flank_seg_cn + actual_cur_seg_cn, data = filter(lm2_modelDF, abs(rd_ci_width - vaf_ci_width) <= 0.05))
summary(someDF2)
```